\chapter*{Why Linear Algebra?}
It is with very little exaggeration that I say that linear algebra is the singular most important area of mathematics. Our intuition for this subject rests largely on how we as humans naturally think about space, transformation, and relationships. Without prompt, we naturally organize data in arrays, think about directions and dimensions, and have well-developed intuition around addition and scaling, which are the fundamental operations of linear algebra. \\

What makes linear algebra so powerful is its remarkable ability to be both a highly effective practical tool and an interesting theoretical subject to study for its own sake. Systems of linear equations appear everywhere, from balancing chemical reactions to modeling physical phenomena to training neural networks. All this while the theoretical side is one of the best ways to introduce new mathematicians to proofs and its mastery and applications often illuminate other areas of math, like our understanding of symmetry (through representation theory) or spaces (through homology).  \\

Linear algebra bridges the concrete and the abstract, and through crossing this bridge many rewards await. A matrix can simultaneously represent a database of information, a system of equations, a geometric transformation, or an abstract linear operator. Much like life, a matrix is what you make it. This versatility explains linear algebra's ubiquity across quantitative fields: computer graphics, quantum mechanics, machine learning, optimization, statistics, differential equations, and far beyond. Your ability to understand applications will provide manifold examples to test your theoretical knowledge, while your mastery of theory helps you make deeper deductions in applications. Each side enriches the other.\\

This book will honor the canonical way of introducing linear algebra, recognizing that this pedagogical structure exists for good reason. The first-year course is mostly calculation-based and is meant for you to get comfortable with performing concrete computations. This computational fluency builds the intuition to appreciate the abstract theory that follows. The second-year course is more abstract, introducing vector spaces axiomatically, exploring the structure of linear transformations, and proving the fundamental theorems that explain why the calculations work. Students who rush to abstraction without computational experience often struggle to develop geometric intuition, while those who master only calculations miss the unifying principles that make linear algebra so powerful. By working through both stages thoughtfully, you'll develop both the technical facility to solve problems and the conceptual framework to understand the deeper structure of the subject.

\chapter{Vectors}
\input{content/algebra/first-year-linear-algebra/1-vectors/first-year-linear-algebra-1-vectors}

\chapter{Solving Linear Equations}
\input{content/algebra/first-year-linear-algebra/2-solving-linear-equations/first-year-linear-algebra-2-solving-linear-equations}

\chapter{Subspaces}
\input{content/algebra/first-year-linear-algebra/3-subspaces/first-year-linear-algebra-3-subspace}

\chapter{The Einstein Summation Convention}
\input{content/algebra/first-year-linear-algebra/13-einstein-summation/first-year-linear-algebra-13-the-einstein-summation-convention}