\chapter*{Why Topology?}
It's 6th period geometry. Your teacher, yet un-assaulted by chalk dust (but give it time), is talking about when we know two triangles are congruent. SSS, SAS, ASA... wait what? Doesn't she mean equal? Why are we just making up new words for equality? Is Big Math™ just making stuff up for the sake of confusion? \\

You're picking up on a subtle point. Two triangles might \emph{seem} equal if they have the exact same side lengths and angles but they aren't for what seems to be a very nit-picky but important reason, they don't occupy the same points in space. However, we have an easy fix. We can find a map called an \vocab{isometry} that maps the set of points that make one triangle onto the set of points that make up the other. Under this rephrasing, the SSS, SAS, ASA conditions are sufficient conditions that allow us to declare the existence of an isometry that maps one triangle onto another without having to explicitly find one. \\

The idea of finding a "useful" class of maps and conditions that guarantee their existence cannot possibly be more prevalent than it is in topology. However, isometries, while useful, are too rigid, they preserve too much structure. That is why, in topology, we begin with the most general class of functions that preserve spatial (topological) information called \vocab{homeomorphisms}. We will discuss conditions that guarantee their existence and spatial information that they preserve. \\ 

While we will (usually) just stick to homeomorphisms in this part, it is important to mention that homeomorphisms are \emph{too} general. A complete classification of spaces up to homeomorphism will elude us as long as Sisyphus remains at his task. So, in practice, we impose additional structure on topological spaces and restrict to maps that preserve this structure, allowing us to detect much finer invariants.\\ 

While homeomorphisms are the maps, the primary actors for which homeomorphisms act on are \vocab{topological spaces}, which are supposed to capture the notion of closeness. It has taken mathematicians many years to pick out the right abstractions of closeness. A primitive approach might be "two points are close if the distance between them is small." But that just kicks the can down the road. "The distance between two points is small if they are close together." The stroke of genius here is that closeness is a contextual property. If I give you one point, you cannot tell me anything about closeness or farness. This is the first hint that sets are the natural tool to talk about closeness. But what \emph{types} of sets? If I just hand you the full collection of points, there just isn't enough information to construct a notion of closeness. We need to pick the \emph{right} sets. \\ 

Let's see how sets can encode closeness through a game. Suppose that we play a game of darts but the bar hasn't ordered the dartboard. Like the hooligans we are, we proceed anyways. We mark a bulls-eye on the blank wall and each throw our dart a single round before the bar owner gets mad at us. Before we are kicked out of the bar, we must determine a winner. The bar owner obliges us and hands us a bunch of empty cardboard in which we can cut out circles. We make a big circle with the first cardboard and place the center on our imagined bullseye. It covers the two holes in the wall we made so we conclude that our darts landed in the interior of our circle. We make a second circle but this time it is far too small. Both holes we made lie in the exterior of our circle. The third circle is our Goldilocks circle, we see that the hole you made lies in the interior and the hole I made lies in the exterior. You win again (how do you keep winning?). Notice that we arrived at a notion of closeness \emph{without measuring or even assigning a number to anything}. Each circle defines an \vocab{open set}: the collection of all points that would fall strictly inside it. By declaring which sets are "open," we encode the information about closeness we need.

\chapter{Topological Spaces and Continuous Functions}
\input{content/geometry-and-topology/point-set-topology/topological-spaces-and-continuous-functions/point-set-topology-1-topological-spaces-and-continuous-functions}

\chapter{Connectedness and Compactness}
\input{content/geometry-and-topology/point-set-topology/connectedness-and-compactness/point-set-topology-2-connectedness-and-compactness}
